On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-12-03 paper.tex
Rewriting and corrections

For $\len \rightarrow \infty$, the average number of visited tiles $\funta$ has the following asymptotic slope, as stems from \eqref{eq: funta, form}:
\begin{equation}
\label{eq: asympt slope ave}
\lim_{\len \rightarrow \infty} \frac{\funta(\len)}{\len} = \frac 2 \pi \left(\frac 1 a + \frac 1 b \right).
\end{equation}
On the other hand, the maximum number of visited tiles $\funt$ has an asymptotic slope given by \eqref{eq: asympt slope}. It is interesting to consider the ratio of these values, which also gives the asymptotic ratio of average and maximum numbers of visited tiles. This depends only on $a/b$:
\begin{equation}
\label{eq: ras}
\ras(a/b) = \lim_{\len \rightarrow \infty}
\frac { \funta(\len) } { \funt(\len) }
= \frac
{ \lim_{\len \rightarrow \infty} \funta(\len)/\len }
{ \lim_{\len \rightarrow \infty} \funt(\len)/\len }
= \frac{2\left(1 + a/b\right)}{\pi\sqrt{1 + \left(a/b\right)^2}}.
\end{equation}
The function $\ras$ satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as can be seen in Figure~\ref{fig: ratio asympt slopes}. It is unimodal with maximum value $2\sqrt 2/\pi \approx 0.9003$ for $a/b=1$, and its limit when $a/b$ tends to $0$ or $\infty$ is $2/\pi \approx 0.6366$. This implies that for $\len$ large the average number of tiles $\funta(\len)$ cannot very small compared with the maximum number of tiles $\funt(\len)$. For example, $\funta(\len)$ asymptotically exceeds $0.8\funt(\len)$ for $1/3 \leq a/b \leq 3$.

@@

For $u, v \in \mathbb N$, $\genvar \in \mathbb R^+$, let
\begin{equation}
\label{eq: g def}
g(\genvar, u, v) = \int_{\arcsin \frac u {2r}}^{\arccos \frac u {2r}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
\end{equation}
Using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$, the integral in \eqref{eq: g def} is computed as
\begin{equation}
\label{eq: g fin}
\begin{split}
g(\genvar, u, v) &= \frac 1 {2\pi} \biggl(
\left(\arccos\frac{u}{2\genvar}-\arcsin\frac{v}{2\genvar}\right) u v + 2\genvar^2 + \frac{u^2} 2 + \frac{v^2} 2 \\
& \quad  - u \sqrt{4\genvar^2-v^2} - v \sqrt{4\genvar^2-u^2} \biggr).
\end{split}
\end{equation}

>>>>

For $\len \rightarrow \infty$, the average number of visited tiles $\funta(\len)$ has the following asymptotic slope, as stems from \eqref{eq: funta, form}:
\begin{equation}
\label{eq: asympt slope ave}
\lim_{\len \rightarrow \infty} \frac{\funta(\len)}{\len} = \frac 2 \pi \left(\frac 1 a + \frac 1 b \right).
\end{equation}
On the other hand, the maximum number of visited tiles $\funt(\len)$ has an asymptotic slope given by \eqref{eq: asympt slope}. It is interesting to consider the ratio of these values, which also gives the asymptotic ratio of average and maximum numbers of visited tiles. This depends only on $a/b$, and is thus expressed as a function $\ras: \mathbb R^+ \to (0,1)$:
\begin{equation}
\label{eq: ras}
\ras(a/b) = \lim_{\len \rightarrow \infty}
\frac { \funta(\len) } { \funt(\len) }
= \frac
{ \lim_{\len \rightarrow \infty} (\funta(\len)/\len) }
{ \lim_{\len \rightarrow \infty} (\funt(\len)/\len) }
= \frac{2\left(1 + a/b\right)}{\pi\sqrt{1 + \left(a/b\right)^2}}.
\end{equation}
The function $\ras$ satisfies $\ras(x) = \ras(1/x)$, and is therefore symmetric in logarithmic scale, as can be observed in Figure~\ref{fig: ratio asympt slopes}. It is unimodal with maximum value $2\sqrt 2/\pi \approx 0.9003$ for $a/b=1$, and its limit when $a/b$ tends to $0$ or $\infty$ is $2/\pi \approx 0.6366$. This implies that for $\len$ large the average number of tiles $\funta(\len)$ cannot very small compared with the maximum number of tiles $\funt(\len)$. For example, $\funta(\len)$ asymptotically exceeds $0.8\funt(\len)$ for $1/3 \leq a/b \leq 3$.

@@

For $u, v \in \mathbb N$, $\genvar \in \mathbb R^+$, let
\begin{equation}
\label{eq: g def}
g(\genvar, u, v) = \frac 2 \pi \int_{\arcsin \frac v {2\genvar}}^{\arccos \frac u {2\genvar}} \left( \genvar \sin \theta - \frac u 2 \right) \left( \genvar \cos \theta - \frac v 2 \right) \diff \theta.
\end{equation}
Computing the integral in \eqref{eq: g def} and using the identities $\cos^2 \alpha = 1-\sin^2 \alpha$ and $\cos(2\alpha) = 2\cos^2\alpha-1 = 1 - 2\sin^2\alpha$,
\begin{equation}
\label{eq: g fin}
\begin{split}
g(\genvar, u, v) &= \frac 1 {2\pi} \biggl(
\left(\arccos\frac{u}{2\genvar}-\arcsin\frac{v}{2\genvar}\right) u v + 2\genvar^2 + \frac{u^2} 2 + \frac{v^2} 2 \\
& \quad  - u \sqrt{4\genvar^2-v^2} - v \sqrt{4\genvar^2-u^2} \biggr).
\end{split}
\end{equation}
